=======
title: "Research Reproducibility: Project 1"
author: "Ibrahimprogramz"
date: "September 29, 2016"
output: html_document
=======

##Setting the global parameters: All codes and its results should appear
```{r global echo}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing data
```{r loading and preprocessing data}
activity<-read.csv("activity.csv")
activity$date<-as.Date(as.character(activity$date))
```

##What is mean total number of steps taken per day?

I loaded the library dplyr in but I have hidden the output as it makes the document looks messy

```{r loading library, include=FALSE}
library(ggplot2)
library(dplyr)
library(chron)
library(lattice)
```

```{r steps means}

#grouping by days to calculate each day and storing into a new dataset to preserve the original
activityday<-group_by(activity, date)
#creating a new dataset containing the means of the grouped days
daytotal<-summarize(activityday,ssum=sum(steps))

#now drawing the result
#drawing the means of each day and breaks is used to count the number of days so that each day mean is on a bar
hist(daytotal$ssum, main = "Steps means to each day from 1/10/2012 to 30/11/2012", xlab = "means of steps for each day", ylab="days", breaks = length(unique(activity$date)), col="blue")
#calculating the means and median of steps for the totals of days
meanMedian<-summarize(daytotal, mean=mean(ssum, na.rm = T), median=median(ssum, na.rm = T))
#printing the mean and median
print(meanMedian)
```

##What is the average daily activity pattern?

```{r activity pattern}
interval<-group_by(activity, interval)
intervalStepMeans<-summarize(interval, stepsmean=mean(steps, na.rm = T))

ggplot() + geom_line(data= intervalStepMeans, aes(x=interval, y=stepsmean), size=.6, color="orange") + ggtitle("Timeseries plot of the steps throuout 5-min day intervals")
indexMax<-which(intervalStepMeans$stepsmean==max(intervalStepMeans$stepsmean))
#The max steps mean is at the interval:
intervalStepMeans[indexMax,1]
```

##Imputing missing values

```{r}
#number of missing values
sum(is.na(activity$steps))

#the strategy is to replace the missing values with day means

#calculating the means without NAs
daytotal1<-summarise(activityday, mean=mean(steps))

activity1<-activity


j<-1
for(i in 1:17568)
{
if(activity1$interval[i]==2355) 
{
  j<-j+1
}
if(is.na(activity1$steps[i])) 
{ 
  activity1$steps[i]<-daytotal1$mean[j]
}

}

activity2<-group_by(activity1, date)
activity2<-summarize(activity2, steps=sum(steps))
hist(activity2$steps, main="total number of steps taken each day",xlab = "Steps",ylab = "Days" ,breaks = 30, col="green")

#the mean and the median
mean(activity2$steps, na.rm = T)

median(activity2$steps, na.rm = T)

#Do these values differ from the estimates from the first part of the assignment? YES
mean(daytotal$ssum, na.rm = T)
median(daytotal$ssum, na.rm = T)

#What is the impact of imputing missing data on the estimates of the total daily number of steps? yes and the difference is:
sum(activity$steps, na.rm = T)-sum(activity1$steps, na.rm = T)
```

##Are there differences in activity patterns between weekdays and weekends?

``` {r}
activity1$day<-weekdays(activity1$date)
#determining which is which regarding weekday and weeek ends
for(i in 1:17568)
{
  if(is.weekend(activity1$date[i]))
   { 
    activity1$day[i]<-"weekend"}
  else
    activity1$day[i]<-"weekday"
}
activity1Weekend<-subset(activity1, day=="weekend")
activity1Weekday<-subset(activity1, day=="weekday")

#weekday data subsetting
activity1IntervalWeekday<-group_by(activity1Weekday, interval)
activity1IntervalStepsWeekday<-summarize(activity1IntervalWeekday, smean=mean(steps, na.rm = T))

#weekend data subsetting
activity1IntervalWeekend<-group_by(activity1Weekend, interval)
activity1IntervalStepsWeekend<-summarize(activity1IntervalWeekend, smean=mean(steps, na.rm = T))



#drawing the data after its was subsetted into weekends and weekdays

ggplot()+ geom_line(data = activity1IntervalStepsWeekend, aes(x=interval, y=smean), col="blue") +ggtitle("Intervals VS Mean of Steps During Weekends") 

ggplot()+ geom_line(data = activity1IntervalStepsWeekday, aes(x=interval, y=smean), col="red") +ggtitle("Intervals VS Mean of Steps During Weekdays")
```
>>>>>>> Project 1
